---
title: "Tidy cell count data"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)

cells <- read_csv(here("data/cellcount/round1/counts_combined.csv")) # copy pasted the summary data into a single spreadsheet, made seperate docs to periodically save
single_cells_1 <- read_csv(here("data/cellcount/round1/individual_cell_stats.csv"))

glimpse(cells)
```

```{r}
library(janitor)

cells2 <- cells %>% 
  clean_names() %>% 
  separate(slice, c("sample_id", "rep"), sep = "_") %>% 
  mutate(rep = rep %>% str_sub(1,2)) %>% 
  rownames_to_column() %>% 
  group_by(sample_id, rep) %>% 
  # take the last rep (when I had to manually edit the image eg erase non-cell things that were counted there were multiple reps, but the last one is correct)
  slice_min(order_by = rowname) %>% 
  ungroup() %>% 
  select(-mean, -rowname) %>% 
  mutate(fraction_cell_cov = percent_area/100, # convert from percent to fraction (easier to do math with)
         .keep = "unused") %>% 
  # rename for clarity
  rename(num_cells = count,
         total_cell_area_px = total_area,
         mean_area_px = average_size)
cells2
```


# single cell areas
```{r}
single_cells_1 %>% 
  ggplot(aes(Area)) +
  geom_histogram()
```
```{r}
cells2 %>% 
  mutate(image_area_px = total_cell_area_px/fraction_cell_cov ) %>% 
  ggplot(aes(image_area_px)) +
  geom_histogram() + 
  ggtitle("number of pixels per image")
# around 3e6 pixels per image
```



```{r}
sc_area <- single_cells_1 %>% 
  filter(Area<1e5) %>% 
  mutate(area_mm2 = (Area /3e6) *2, # convert pixels to um2, assuming each image has 3e6 pixels per image covering 0.5 mm2
         area_um2 = area_mm2 * (1000 * 1000),
         diam_um = sqrt(area_um2/pi)*2) #A = pi*r^2 -> r = sqrt(a/pi)
sc_area %>% 
  ggplot(aes(diam_um)) +
  geom_histogram()
```

```{r}
sc_area %>% 
  summarise(mean = mean(diam_um),
            sd = sd(diam_um))
# 95% between 2 SDs
# 10-70 diam um
```
Differences between samples?
Need the cell numbers to match to samples...
```{r}

```



Bring in the 0's
```{r}
library(fs)

zeros <- dir_ls(here("photos/cell_count_photos/zero_cells")) %>% 
  basename() %>% 
  str_sub(1,11) %>% 
  as_tibble() %>% 
  separate(value, c("sample_id", "rep"), sep = "_")

cells3 <- cells2 %>%
  bind_rows(zeros) %>%
  replace_na(list(num_cells = 0, total_cell_area_px = 0, fraction_cell_cov = 0, total_cell_area_mm2 = 0)) %>% 
  arrange(sample_id, rep)

# sanity check
cells3 %>% 
  group_by(sample_id) %>% 
  summarize(n = n()) %>% 
  arrange(n)
```

Missing one count for whi21.10.
All other samples have the expected 16 counts

Sanity check, get distribution of mean cell area
```{r}
cells3 %>% 
  ggplot(aes(mean_area_px)) +
  geom_histogram()
```

Check on outlier

```{r}
cells3 %>% 
  filter(mean_area_px > 3e5)
```

Check the photos. There are no cells in either photo. Set to 0

```{r}
cells3 <- cells3 %>% 
  rownames_to_column()
cells3 %>% 
  filter((sample_id == "tri21.12" & rep=="05") |
         (sample_id == "tri21.12" & rep == "15"))
```

Some issue with a duplicate?

```{r}
cells4 <- cells3 %>% 
  filter(rowname != 271) %>% # remove the non zero one
  # set the erroneous one to 0
  mutate(num_cells = if_else(rowname==261, 0, num_cells),
         total_cell_area_px = if_else(rowname==261, 0, total_cell_area_px),
         mean_area_px = if_else(rowname==261, NA_integer_, mean_area_px %>% as.integer()),
         fraction_cell_cov = if_else(rowname==261, 0, fraction_cell_cov))
# sanity check
cells4 %>% 
  filter(rowname==261)
```


```{r}
# check the remaining
cells4 %>% 
  ggplot(aes(mean_area_px)) +
  geom_histogram()
```

Check smaller other outlier


```{r}
cells4 %>% 
  filter(mean_area_px > 10000)
```

The image looks like a zero

```{r}
cells5 <- cells4 %>% 
    mutate(num_cells = if_else(rowname==16, 0, num_cells),
         total_cell_area_px = if_else(rowname==16, 0, total_cell_area_px),
         mean_area_px = if_else(rowname==16, NA_integer_, mean_area_px %>% as.integer()),
         fraction_cell_cov = if_else(rowname==16, 0, fraction_cell_cov))
cells5 %>% filter(rowname==16)
```


```{r}
# check the remaining
cells5 %>% 
  ggplot(aes(mean_area_px)) +
  geom_histogram()
```

Check the final outlier

```{r}
cells5 %>% 
  filter(mean_area_px>7000)
```

Just a clump, OK



```{r}
cells5 %>% 
  ggplot(aes(x = fraction_cell_cov)) +
  geom_histogram()
```

Check outlier

```{r}
cells5 %>% 
  filter(fraction_cell_cov > 0.03)
```

This is a zero. Not sure how the program counted 48?

```{r}
cells6 <- cells5 %>% 
    mutate(num_cells = if_else(rowname==264, 0, num_cells),
         total_cell_area_px = if_else(rowname==264, 0, total_cell_area_px),
         mean_area_px = if_else(rowname==264, NA_integer_, mean_area_px %>% as.integer()),
         fraction_cell_cov = if_else(rowname==264, 0, fraction_cell_cov))
```


```{r}
cells6 %>% 
  ggplot(aes(x = fraction_cell_cov)) +
  geom_histogram()
```

```{r}
cells6 %>% 
  arrange(-fraction_cell_cov)
```

Seems to match the photos, OK





# Compute final stats per sample

```{r}
cells7 <- cells6 %>% 
  # reps are grouped by pairs for each grid, and each 4 grids belong to a statistical replicate (an aliquot from the same sample)
  mutate(rep = as.numeric(rep),
         grid = case_when(rep %in% 1:2 ~ "a",
                          rep %in% 3:4 ~ "b",
                          rep %in% 5:6 ~ "c",
                          rep %in% 7:8 ~ "d",
                          rep %in% 9:10 ~ "e",
                          rep %in% 11:12 ~ "f",
                          rep %in% 13:14 ~ "g",
                          rep %in% 15:16 ~ "h"),
         aliquot = case_when(grid %in% letters[1:4] ~ "A",
                             grid %in% letters[5:8] ~ "B")) %>% 
  # combine the two halves
  group_by(sample_id, grid, aliquot) %>% 
  summarize(num_cells = sum(num_cells), # add the two half grids together
            total_cell_area_px = sum(total_cell_area_px), # add the two half areas together
            mean_area_px = mean(mean_area_px, na.rm = TRUE), # mean cell size for the two halves (some NA's, must set na.rm true)
            fraction_cell_cov = mean(fraction_cell_cov)) %>% # mean fraction cell coverage for the two halves 
  ungroup()

cells7
```



Since each set of 4 grids is a psuedo-replicate, add them together for each aliqout
ie compute cell area per 0.4 ul

Also convert pixels to real units

Compute:
- cells per ul
- total cell area per ul
- mean cell area


```{r}
cells8 <- cells7 %>% 
  # doing the algebra to convert units:
  # fraction  = cell area / total area -> cell area = fraction * total area
  # total area is 1 mm2...
  # so cell area in mm2 = fraction 
  rename(cell_area_mm2 = fraction_cell_cov) %>% # per 1x1 mm grid
  # compute stats per 0.4 ul (the total of 4 grids, there's 0.1 ul per grid)
  group_by(sample_id, aliquot) %>% 
  summarise(cells_per_ul = sum(num_cells) / 0.4, 
            # convert units mm2 per ul to um2 per ul
            cell_um2_per_ul = (sum(cell_area_mm2) *(1000*1000)) / 0.4) %>% 
  ungroup() %>% 
  # use the result to compute mean cell area, the vol units cancel
  mutate(mean_cell_area_um2 = cell_um2_per_ul/cells_per_ul) %>% 
  select(sample_id, aliquot, cells_per_ul, cell_um2_per_ul, mean_cell_area_um2)

cells8 %>% 
  arrange(sample_id)
```

The individual cell areas are reasonable, e.g. a cell 5 um radius has area of ~80 um2, of 10 um radius would have radius of 300um2

The total cell areas are reasonable, if we divide by 10 we have on the order of magnitude of 1,000 - 10,000 um2 per grid. The grid is 1,000,000 um2 total, so we see a range between .1 and 1 percent coverage of the cells. 

The cells per ul is reasonable, if we divide by 10 we have the n cells per grid. 


```{r}
cells8 %>% 
  write_csv(here("data/cellcount/clean_cell.csv"))
```

