---
title: "combine final cell count data, compute stats"
output: rmarkdown::github_document
---

Using the cookie samples collected w Scott 2021

```{r libraries}
library(here)
library(tidyverse)
```

# Read in data
```{r}
cellcount_1 <- read_csv(here("data/cellcount/round1/clean_cell.csv"))
cellcount_2 <- read_csv(here("data/cellcount/round2/clean_cell_recount.csv"))
raw_epi <- read_csv(here("data/field_sample_meta/tidy_epicollect.csv"))
```

```{r}
glimpse(cellcount_1)
```
```{r}
glimpse(raw_epi)
```
# wrangle

## convert snow weight to liquid volume 

```{r}
epi <- raw_epi %>% 
  filter(container_type=="ziploc") %>% # only did cell counts for these samples
  mutate(total_sample_water_vol_ul = weight_g * 1000, .keep = "unused")  %>% # 1 g h2o = 1 mL = 1000 ul
  distinct(sample_id, total_sample_water_vol_ul) # remove duplicate entries
```


## join datasets

```{r}
joined <- cellcount_1 %>% 
  bind_rows(cellcount_2) %>%
  left_join(epi)

joined
```

## Calculate fraction red cover

1. scale up area to entire sample

the ratio should remain the same
- A1/V1 = A2/V2
- cookie cell area/cookie liquid volume = hemocytometer cell area / hemocytometer grid volume
or solving for cookie cell area:
- cookie cell area = (hemocytometer cell area / hemocytometer grid volume) * cookie liquid volume

2. Then express cookie total cell area as a fraction of the cookie area

- frac_red = cookie cell area/cookie area

3. putting the two together

- frac_red =  ((hemocytometer cell area / hemocytometer grid volume) * cookie liquid volume ) / cookie area
or since our counts are already per hemocytometer grid vol:
- frac_red =  (cell_um2_per_ul * cookie liquid volume ) / cookie area

```{r}
cookie_area_um2 <- pi*(50 * 1000)^2 # PVC sampler radius = 5cm = 50 mm = 50,000 um
cookie_area_sq_m <- pi*0.05^2

joined_frac <- joined %>% 
  mutate(frac_red = (cell_um2_per_ul * total_sample_water_vol_ul) / cookie_area_um2,
         cells_per_sq_m = (cells_per_ul * total_sample_water_vol_ul) /cookie_area_sq_m,
         # convert the units to cm2 per l for comparability with Ganey 2017. 
         cell_cm2_per_l = cell_um2_per_ul * 0.01) # 1 um2/uL * 1cm2/1e8um2 * 1e6 uL/1L = 0.01 cm2/L

joined_frac %>% 
  arrange(-frac_red)
```

Check that the units are comparable to Ganey et al
```{r}
joined_frac %>% 
  select(cell_cm2_per_l) 
```
Ganey range is 100-300 cm2/L

# EDA

## histograms
```{r}
cell_stats <- joined_frac %>% 
  select(-aliquot) %>% 
  # group by sample
  group_by(sample_id) %>% 
  summarize(across(everything(), list(mean = mean, sd = sd), .names = "{.col}.{.fn}")) %>% 
  ungroup() %>% 
  pivot_longer(-sample_id) %>% 
  separate(name, c("var", "stat"), sep = "\\.") 

cell_stats %>% 
  filter(stat == "mean") %>% 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(vars(var), scales = "free") +
  ggtitle("histograms of cell count data")

```
## Mean cell area

assume cells are spherical
```{r}
joined_frac %>% 
  mutate(mean_rad_um = sqrt(mean_cell_area_um2/pi), # a=pi*r^2 -> r = sqrt(a/pi)
         mean_vol_um3 = (4/3)*pi*mean_rad_um ) %>% # v=(4/3)pi*r 
  ggplot(aes(mean_rad_um)) +
  geom_histogram()
```

## Outlier check

```{r}
joined_frac %>% 
  filter(cells_per_ul > 250)
```

## histograms of SD
```{r}
cell_stats %>% 
  filter(stat == "sd") %>% 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(vars(var), scales = "free") +
  ggtitle("SD histograms of cell count data")
```

Most counts have low SD which is heartening. One high outlier

```{r}
cell_stats %>% 
  filter(var == "cells_per_ul", stat == "sd", value > 40)
```

tri21.03 Aliquot A got some clumps, hence the high SD


```{r}
cell_stats_wide <-
  cell_stats %>% 
  unite(name, var:stat, sep = "_") %>% 
  pivot_wider(sample_id)

cell_stats_wide %>% 
  arrange(-frac_red_sd) %>% 
  select(sample_id, frac_red_sd)
```

Convert to cells per mL for comparison with Painter 2001
```{r}
cell_stats_wide %>% 
  mutate(cells_per_ml_mean = cells_per_ul_mean * 1000) %>%
  ggplot(aes(y=cells_per_ml_mean, x = 0)) +
  geom_point() 
```
```{r}
cell_stats_wide %>% 
  mutate(cells_per_ml_mean = cells_per_ul_mean * 1000) %>%
  summarise(mean = mean(cells_per_ml_mean))
```



# Write csv

```{r}
joined_frac %>%
  write_csv(here("data/cellcount/final_cell.csv"))
```



