---
title: "Snow algae abundance regression"
output: rmarkdown::github_document
---

```{r}
library(tidyverse)
library(here)
library(fs)
library(janitor)
library(GGally) 
library(broom) 
library(grid)
library(ggpubr)
# library(brms)
```

# read in data

```{r}
cellcount <- read_csv(here("data/cellcount/final_cell.csv"), col_types = cols()) 
biomass <- read_csv(here("data/biomass/tidy_biomass.csv"), col_types = cols()) 
color_raw <- read_csv(here("data/field_sample_meta/final_sample_list.csv"))

# convolved sample scans
as_s2 <- read_csv(here("data/radiometer/as_s2.csv"), col_types = cols())
as_l8 <- read_csv(here("data/radiometer/as_l8.csv"), col_types = cols())
as_terra <- read_csv(here("data/radiometer/as_terra.csv"), col_types = cols())
as_planet <- read_csv(here("data/radiometer/as_planet.csv"), col_types = cols())
```


# tidy the cell count data

```{r}
cell_select <- cellcount %>% 
  select(sample_id, frac_red, cell_cm2_per_l, cells_per_sq_m) %>% # keep cm^2/L for comparison w Ganey 2017
  # convert units
  mutate(bil_cells_per_sq_m = cells_per_sq_m/1e9, # convert to billion cells
         cell_area_percent = frac_red * 100,
         .keep = "unused")
```



# Compute RGND

Red green normalized difference (RGND; referred to as NDI in Ganey 2017; SDI in DiMauro 2015)

```{r}
# a function to compute the normalized difference
nd <- function(a,b){ (a-b)/(a+b) } 

compute_rgnd <- function(data, green, red){
  green <- enquo(green)
  red <- enquo(red)
  
  data %>% 
    mutate(rgnd = nd(!!red, !!green)) %>% 
    select(scan_id, sample_id, rgnd)
}

# compute indices for convolved radiometer data 
s2_rgnd <- as_s2 %>% 
    compute_rgnd(green = b3, red = b4)
l8_rgnd <- as_l8 %>% 
    compute_rgnd( green = band_3, red = band_4)
terra_rgnd <- as_terra %>% 
    compute_rgnd(green = RSR_555, red = RSR_645)
```

## Cross-platform comparison

```{r}
# make a new tibble, with added column specifying the satellite platform name
rgnd_comparison <- add_column(s2_rgnd, platform = "s2") %>% 
  bind_rows(add_column(l8_rgnd, platform = "l8")) %>% 
  bind_rows(add_column(terra_rgnd, platform = "terra")) %>% 
  pivot_wider(names_from = platform, values_from = rgnd)
  

xa <- rgnd_comparison %>% 
  ggplot(aes(y =s2, x = terra)) +
  geom_point(alpha = 0.7)+
  theme_minimal() +
  labs(y = "Sentinel-2 RGND", x = "MODIS  RGND", tag = "A")

xb <- rgnd_comparison %>% 
  ggplot(aes(y =s2, x = l8)) +
  geom_point(alpha = 0.7)+
  theme_minimal() +
  labs(y = "Sentinel-2 RGND", x = "Landsat-8 RGND", tag = "B")

xc <- rgnd_comparison %>% 
  ggplot(aes(x =terra, y = l8)) +
  geom_point(alpha = 0.7)+
  theme_minimal() +
  labs(x = "MODIS RGND", y = "Landsat-8 RGND", tag = "C")

ggarrange(xa, xb, xc)
```
```{r}
ggsave(here("figs/rsr_rgnd_pairs.png"))
```



```{r}
rgnd_comparison_long <- rgnd_comparison %>% 
  pivot_longer(s2:terra)

rgnd_comparison_long %>% 
  ggplot(aes(x = name, y = value)) +
  geom_boxplot() +  
  geom_jitter(width = 0.2, alpha = 0.2) +
  labs(title= "Field radiometer snow algae sample scan RGND convolution comparison", x = "Satellite platform", y = "RGND")
```
ANOVA comparison of the above boxplot

```{r}
compare_mod <- lm(value ~ name, data = rgnd_comparison_long)
summary(compare_mod)
```
No significant difference in RGND computed using different satellite RSRs. 

I will now arbitrarily choose S2 RGND for the remaining analysis. 



## QC for duplicate scans discrepancies

This is similar to the duplicate scan comparison in convolve_sample_scans.R but with RGND instead of spectral albedo 


Compare the duplicate scan values convolved to RGND
```{r}
color <- color_raw %>%
  mutate(site = sample_id %>% str_sub(1,3),
         num = sample_id %>% parse_number(),
         sample_id = paste0(site, "21.", as.character(num) %>% str_pad(2,pad = "0"))) %>% 
  select(sample_id, type)
write_csv(color, here("data/field_sample_meta/final_sample_list_shortname.csv"))

s2_rgnd %>% 
  left_join(color) %>% 
  ggplot(aes(y = rgnd, x= sample_id %>% fct_reorder(rgnd), color = type)) +
  geom_point() +
  labs(y = "RGND", x = "Sample ID") +
  coord_flip() +
  scale_color_manual(values = c("black","red", "purple", "grey90"))
```


```{r}
s2_rgnd %>% 
  filter(sample_id=="whi21.10")
```



Take mean value for each sample

```{r}
rgnd_final <- s2_rgnd %>% 
  group_by(sample_id) %>% 
  summarise(rgnd_mean = mean(rgnd), rgnd_sd = sd(rgnd))
```


# Plots


```{r}
# combine the data into a single data frame

# first make a long version, keeping the SD for plotting with crossbars
cell_long <- cell_select %>%  
  pivot_longer(-sample_id) %>% 
  group_by(sample_id, name) %>% 
  summarize(mean = mean(value), sd = sd(value)) %>% 
  ungroup()

# save this for use in the albedo regression
cell_long %>% write_csv(here("data/cellcount/cell_long.csv"))

mass_long <- biomass %>% 
  pivot_longer(-sample_id, values_to = "mean") %>% # common var name for joining
  drop_na(mean) # missing IC for a few samples

cell_mass_long <- cell_long %>% 
  bind_rows(mass_long)


joined_long <- rgnd_final %>% 
  inner_join(cell_mass_long, by = "sample_id") %>% # use inner join to drop scans with no corresponding lab values
  # shorten sample IDs for plotting
  mutate(sample_id = sample_id %>% str_remove("21.0"),# Remove the 21. from the sample IDs
       sample_id = sample_id %>% str_remove("21."))

# also make a simpler wide version, with no SD 
joined_wide <- joined_long %>% 
  select(-sd, -rgnd_sd) %>% 
  pivot_wider(names_from = "name", values_from = "mean") %>% 
  rename(rgnd = rgnd_mean)
```



## pairs plot (all vars)

```{r}
pairs_plot <- joined_wide %>%
  select(-sample_id) %>% 
  ggpairs()
# view this on full screen, hard to see within markdown doc
```

```{r}
joined_wide %>% 
  arrange(-mass_g_per_m2)
```



Several samples had outlier high TOC despite having relatively low algae, these samples contained visibly high mineral dust/pollen/dark stuff on surface, likely containing high organics that wouldn't increase the NDVI or RGND. Remove these 

```{r}
dirty_snow_samples <- c("bdw9", "whi10", "tri14")

joined_wide_clean <- joined_wide %>% 
  filter(!(sample_id %in% dirty_snow_samples))
joined_long_clean <- joined_long %>% 
  filter(!(sample_id %in% dirty_snow_samples))
```




There are two outliers in the pairs plot of TOC ~ frac_red, and TOC ~ cellcount plots.... what are they?

## TOC vs algae

```{r}
joined_wide %>% 
  ggplot(aes(y = toc_g_per_m2, x = cell_area_percent)) +
  geom_point() +
  geom_text(aes(label = sample_id), alpha = 0.4)

joined_wide %>% 
  ggplot(aes(y = toc_g_per_m2, x = bil_cells_per_sq_m)) +
  geom_point() +
  geom_text(aes(label = sample_id), alpha = 0.4)
```

tri3 and whi3 are the outliers
Possible explanations: 
- error in measurement, discrepancy between radiometer measurement point and the cookie (see photo).
- cleaner, these had more algae and less biofilm. 

## cell area vs count

```{r}
joined_wide%>% 
  ggplot(aes(y = cell_area_percent, x = bil_cells_per_sq_m)) +
  geom_point() +
  geom_text(aes(label = sample_id), alpha = 0.4)
```
tri3 and whi3 have lower cell area than the others, the cells are just smaller. 


## fig 2: RGND predicts Snow Algae cells, area, TOC, TN


Use the long data for plotting with facets
```{r, fig.width = 3, fig.height = 7}
# use facetting to plot to automatically align labs

# (could do this manually by setting scales, but it dosen't precisely align the axis due to different fonts at edges, etc)
vars_to_plot <- c("bil_cells_per_sq_m", "cell_area_percent", "toc_g_per_m2", "tn_mg_per_m2")

fig2_dat <- joined_long_clean %>%
  filter(name %in% vars_to_plot) %>%
  # order the factors, for plotting
  mutate(name = name %>% fct_relevel(vars_to_plot)) # pull(resp_name) %>% levels() #check fct levels

facet_labs = c(
  bil_cells_per_sq_m = "Algae per area (10^9 cells/m2)",
  cell_area_percent = "Algal % coverage",
  toc_g_per_m2 = "TOC (g/m2)",
  tn_mg_per_m2 = "TN (mg/m2)"
)

f2 <- fig2_dat %>% 
  ggplot(aes(y = mean, x = rgnd_mean)) +
  # plot the OLS regression lines...
  # specify a 0 intercept for snow algae cell count and cell area
  geom_smooth(method = "lm", alpha = 0.6, se = FALSE, color = "lightblue", formula = y ~ 0 + I(x^2),
              data = fig2_dat %>% filter(name %in% c("bil_cells_per_sq_m", "cell_area_percent"))) + 
  # geom_smooth(method = "lm", alpha = 0.6, se = FALSE, color = "green", formula = y ~ x,
  #             data = fig2_dat %>% filter(name %in% c("cell_area_percent"))) +
  # allow y intercept to range freely for TOC and TN plots
  geom_smooth(method = "lm", alpha = 0.6, se = FALSE, color = "lightblue", formula = y ~ x,
              data = fig2_dat %>% filter(!(name %in% c("cell_area_percent", "bil_cells_per_sq_m")))) + 
  geom_point(alpha = 0.4, size = 2) +
  # the x val SD crossbar
  geom_linerange(aes(ymin = mean - sd, 
                      ymax = mean + sd),
                      alpha = 0.4) +
  # the y val SD crossbar
  geom_linerange(aes(xmin = rgnd_mean - rgnd_sd,
                   xmax = rgnd_mean + rgnd_sd),
                alpha = 0.4) +
  facet_grid(rows = vars(name), # to avoid redundant xlabs use _grid, not _wrap
             scales = "free", 
             labeller = labeller(name = facet_labs)) +
  # annotate outliers
  geom_text(aes(label = sample_id), size = 2.5, alpha = 0.5, nudge_x = -0.03,
            data = fig2_dat %>% filter(sample_id %in% c("whi3", "bdw4"))) +
  theme_minimal() +
  theme(axis.title.y = element_blank()) +
  labs(x = "RGND") +
  theme(panel.spacing = unit(2, "lines")) # increase spacing between panels
f2
ggsave(here("figs/figs_v2/3_abundance/biomass.pdf"), width = 3, height = 7) # polish in inkscape
```
The final axis labels should be:
Percent snow algae cell area
Cell abundance (10^9 cells/m^2)
Total organic carbon (g/m^2)
Total nitrogen (mg/m^2)

(add these in the final polish in Inkscape)



# OLS regression

lm models
```{r}
cell_mod <- lm(bil_cells_per_sq_m ~ 0 + I(rgnd^2), data = joined_wide_clean) # set intercept to 0 for simplicity
frac_mod <- lm(cell_area_percent ~ 0 + I(rgnd^2), data = joined_wide_clean) # set intercept to 0 for simplicity
toc_mod <- lm(toc_g_per_m2 ~ rgnd, data = joined_wide_clean)
tn_mod <- lm(tn_mg_per_m2 ~ rgnd, data = joined_wide_clean)
```

## lm coefs
```{r}
coefs <- bind_rows(
  tidy(cell_mod) %>% add_column(response = "cell abundance"),
  tidy(frac_mod) %>% add_column(response = "cell area"),
  tidy(toc_mod) %>% add_column(response = "toc"),
  tidy(tn_mod) %>% add_column(response = "tn")) %>% 
  relocate(response) 
coefs
```


cell abundance= 22 +/- 3.5 RGND^2 (+/- SE)
cell area= 694 +/- 85 RGND^2
TOC= 16 +/- 2.2 RGND
TN= 365 +/- 48 RGND


Export coefficient table for use in biogeography paper
```{r}
coefs %>% 
  write_csv(here("data/abundance_coefs.csv"))
```



Compare without constraint of B0 (y intercept)
```{r}
cell_mod2 <- lm(bil_cells_per_sq_m ~ I(rgnd^2), data = joined_wide_clean)
frac_mod2 <- lm(cell_area_percent ~ I(rgnd^2), data = joined_wide_clean) 
summary(cell_mod2)
summary(frac_mod2)
```

Estimated y intercept within one standard error of zero


## R squared
```{r}
summary(cell_mod)$r.squared
summary(frac_mod)$r.squared

summary(toc_mod)$r.squared
summary(tn_mod)$r.squared
```



## Compare polynomial with single term models 


```{r}

cell_poly_mod <- lm(bil_cells_per_sq_m ~ 0 + I(rgnd^2), data = joined_wide_clean %>% drop_na(rgnd))
frac_poly_mod <- lm(cell_area_percent ~ 0 + I(rgnd^2), data = joined_wide_clean)
toc_poly_mod <- lm(toc_g_per_m2 ~ 0 + poly(rgnd, 2), data = joined_wide_clean %>% drop_na(rgnd))
tn_poly_mod <- lm(tn_mg_per_m2 ~ 0 + poly(rgnd, 2), data = joined_wide_clean %>% drop_na(rgnd))



# compare AIC, choose model for each variable that minimizes AIC
AIC(frac_poly_mod, frac_mod, 
    cell_poly_mod,cell_mod,
    toc_poly_mod,toc_mod,
    tn_poly_mod, tn_mod) %>% 
  rownames_to_column("model") %>% 
  mutate(quadratic = ifelse(model %>% str_detect("poly"), TRUE, FALSE),
         response_variable = model %>% str_remove("_mod") %>% str_remove("_poly")) %>% 
  select(-df, -model) %>% 
  pivot_wider(names_from = response_variable, values_from = AIC)

```

AIC is minimized with the single term model in all cases



## compare w Ganey et al 2017 "algae abundance model" regression

```{r}
joined_wide_clean %>% 
  ggplot(aes(x = rgnd, y = cell_cm2_per_l)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ 0 + x, color = "blue") +
  geom_abline(0, slope = 1449, linetype = "dashed", color="red") +
  theme_minimal() +
  labs(x = "RGND", y = "Snow algae cell area (cm2/L)")
```

```{r}
ggsave(here("figs/compare_vs_ganey.pdf"),width = 7, height = 5)
```

Ganey : A = 1,449 NDI


```{r}
ganey_frac_mod <- lm(cell_cm2_per_l ~ 0 + rgnd, data = joined_wide_clean)
tidy(ganey_frac_mod)
summary(ganey_frac_mod)$r.squared

```
```{r}
1727/1449 # our est is +19% of ganey
1449/1727 # ganey's estimate is 16% lower than ours
```
excluding two highest outliers  
```{r}
ganey_frac_mod2 <- lm(cell_cm2_per_l ~ 0 + rgnd, data = joined_wide_clean %>% 
                       filter(cell_cm2_per_l<800))
tidy(ganey_frac_mod2)
summary(ganey_frac_mod2)$r.squared
```

<!-- # Bayesian regression -->


<!-- Conjugate priors -->

<!-- ```{r} -->
<!-- var1 <- -->
<!-- priors1 <- c(set_prior("normal(3, 0.632)", class = "b", coef = "rgnd")) -->

<!-- cell_mod <- brm(bil_cells_per_sq_m ~ 0 + rgnd, data = joined_wide_clean, seed = 123, priors = priors1) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- priors2 <- c(set_prior("normal(3, 0.632)", class = "b", coef = "age")) -->

<!-- frac_mod <- brm(cell_area_percent ~ 0 + rgnd, data = joined_wide_clean, seed = 123) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- priors3 <- c(set_prior("normal(3, 0.632)", class = "b", coef = "age"), -->
<!--              set_prior("normal(0, 0.316)", class = "b", coef = "age2")) -->
<!-- toc_mod <- brm(toc_g_per_m2 ~ rgnd, data = joined_wide_clean, seed = 123) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- priors4 <- c(set_prior("normal(3, 0.632)", class = "b", coef = "age"), -->
<!--              set_prior("normal(0, 0.316)", class = "b", coef = "age2")) -->
<!-- tn_mod <- brm(tn_mg_per_m2 ~ rgnd, data = joined_wide_clean, seed = 123) -->
<!-- ``` -->




