---
title: "Satellite image comparison"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
library(here)
library(fs)
library(janitor)
library(lubridate)
library(tsibble)
library(splines)
library(zoo)
library(kableExtra)
library(GGally)
```

```{r}
# modis polygon
ts_modis <- dir_ls(here("data/gee_ts/modis_grid_cells"), regexp = ".csv$") %>% 
  map_df(read_csv, id = "filename", col_types = cols())

# bloom polygon
ts_custom <- dir_ls(here("data/gee_ts/custom_polygon"), regexp = ".csv$") %>% 
  map_df(read_csv, id = "filename", col_types = cols())

# glacier polygon
ts_glims <- dir_ls(here("data/gee_ts/glimsGeom"), regexp = ".csv$") %>% 
  map_df(read_csv, id = "filename", col_types = cols())

# histogram of RGND during peak bloom week
hist <- dir_ls(here("data"), regexp = "*hist.csv$") %>% 
  map_df(read_csv, id = "filename", col_types = cols())

# wx
wx <- dir_ls(here("data"), regexp = "*wx.csv$") %>% 
  map_df(read_csv, id = "filename", col_types = cols())
```


# tidy

```{r}
# parse site name, date
my_tidy <- function(df){
  df %>% 
    clean_names() %>% 
    mutate(site = filename %>% basename() %>% str_remove(".csv") %>% str_split_n("_", 1),
           date = mdy(system_time_start))
}

# parse platform name, apply cloudmask
ts_tidy <- function(df){
  df %>% 
     mutate(platform = basename(filename) %>% 
              str_split_n("_", 2) %>% 
              str_remove(".csv"),
         rgnd = if_else(is.na(rgnd), rgnd_cloudmasked, rgnd)) %>% 
  select(platform, site, date, rgnd, albedo)
}


ts_modis_wide <- ts_modis %>% 
  my_tidy() %>% 
  ts_tidy()

ts_custom_wide <- ts_custom %>% 
  my_tidy() %>% 
  ts_tidy()

ts_glims_wide <- ts_glims %>% 
  my_tidy() %>% 
  ts_tidy()

hist_clean <- hist %>% 
  clean_names() %>% 
  mutate(site = basename(filename) %>% str_split_n("_", 1), .keep = "unused")

wx_clean <- wx %>% 
  my_tidy() %>% 
  select(-filename, -system_time_start) %>% 
  pivot_longer(-site:-date, values_drop_na =TRUE) %>% 
  pivot_wider()
  
```





# fig6 plot


```{r}
# make pretty for plotting
plot_dat <- ts_custom_wide %>% 
  filter(platform %in% c("s2", "l8", "terra"),
         # remove sept 8 for catamount
         !(platform=="s2" & date==ymd("2020-09-08"))
         ) %>% 
  mutate(platform = platform %>% 
           fct_recode("Sentinel-2"="s2",
                      "Landsat-8"="l8",
                      "MODIS/Terra"="terra") %>% 
           fct_relevel(c("Sentinel-2", "Landsat-8")),
         site = site %>% fct_relevel("vowell")) 
plot_dat$platform %>% levels()

gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

p1 <- plot_dat %>% 
  ggplot(aes(x = date, y = rgnd, color = platform, shape = platform)) +
  facet_grid(rows = vars(site)) +
  geom_point(alpha =0.7) +
  labs(x = "Date", y = "RGND", color = "Satellite") +
  theme_minimal() +
  geom_hline(yintercept = 0) +
  scale_color_manual(values = gg_color_hue(3)) +
  theme(panel.spacing = unit(2, "lines")) 
p1
```


## sentinel linear interpolation

```{r}
s2_interp <- plot_dat %>% 
  filter(platform=="Sentinel-2", 
         # just July-Aug data
         month(date) %in% c(7,8)) %>% 
  # set 0 on Jul 1 and Sept 1
  bind_rows(tribble(
    ~site, ~date, ~rgnd,
    "cat", ymd("2020-07-01"), 0,
    "cat", ymd("2020-09-01"), 0,
    "vowell", ymd("2020-07-01"), 0,
    "vowell", ymd("2020-09-01"), 0
  )) %>% 
  # convert to tsibble to use handy fill_gaps()
  as_tsibble(key = site) %>%
  # make missing values explicit NA
  fill_gaps() %>% 
  # linear interpolate NA
  mutate(rgnd = zoo::na.approx(rgnd)) %>% 
  as_tibble() %>% 
  # add stuff back in for plotting
  mutate(platform="Sentinel-2",
         site = site %>% fct_relevel("vowell"))
```

## export final plot

```{r}
p1 +
  # geom_line(stat="smooth", method = "loess", data= curve_dat, alpha = 0.5, size = 1, color = "purple")
  geom_line(data= s2_interp, alpha = 0.5, size = 1, color = gg_color_hue(3)[1]) +
  geom_point(alpha = 0.7)

```

```{r}
ggsave(here("figs/satellite_time_series.pdf"), height = 5, width = 7)
```




# Compare satellite RGND w/in 5 day window

## pairs plot
```{r}
date_bins <- plot_dat %>% 
  # bin time into discrete 4 day intervals
  mutate(date_bin = date %>% yday() %>% cut_width(width=5)) %>% 
  group_by(platform, site, date_bin) %>% 
  summarise(rgnd = mean(rgnd)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = "platform", values_from = "rgnd")

# check that cut_width works as expected
date_bins %>% 
  distinct(date_bin) %>% 
  arrange(date_bin)
```

Comparison of mean values for each 5 day time interval
```{r}
date_bins %>%
  select(where(is.numeric)) %>% 
  ggpairs() 
```
## peak bloom boxplot


```{r}
peak_yday <- yday(ymd('2020-07-28'))
# i can see this date bin is (208,212], thats Jul 27 -Jul 30


plot_dat %>% 
  filter(date>ymd("2020-07-26"), date <= ymd("2020-07-30")) %>% 
  ggplot(aes(x = platform, y = rgnd)) +
  geom_boxplot() +
  facet_wrap(vars(site))
```
defining "peak week" as Jul 27 -Jul 30

# algae stats during peak week


## bloom polygon

Apply regression estimates:

cell abundance: 6.95 +/- 3.11 RGND
cell area: 216 +/- 78 RGND
TOC: 16 +/- 6.5 RGND
TN: 365 +/-142 RGND
delta_albedo = 0.9 +/- 0.34 RGND 

```{r}
# from GEE
bloom_polygon_area <- tribble(
  ~site, ~area_m2,
  "vowell", 1684302,
  "cat", 1918716
)
```


```{r}
get_max_s2_rgnd <- function(df){
  df %>% 
  filter(platform == "s2") %>% 
  group_by(site) %>% 
  summarise(rgnd = max(rgnd, na.rm = TRUE))
}
# ts_custom_wide %>% get_max_s2_rgnd()


estimate_bloom_stats <- function(rgnd_df, polygon_area_df){
  rgnd_df %>%
    left_join(polygon_area_df) %>% 
    mutate(
      cell_count = 6.95 * rgnd, # bil cells / m2
      cell_count_ci = 3.11 * rgnd,
      cell_area = 216 * rgnd, # percent cell area
      cell_area_ci = 78 * rgnd,
      toc = 16 * rgnd, # g/m2
      toc_ci = 6.5 * rgnd,
      tn = 365 * rgnd, # mg/m2
      tn_ci = 142 * rgnd,
      delta_albedo = 0.9 * rgnd, # delta albedo
      delta_albedo_ci = 0.34 * rgnd, 
      total_cells = cell_count * area_m2, # bil cells
      total_cells_ci = cell_count_ci * area_m2, 
      total_toc_kg = (toc * area_m2)/1000, # kg
      total_toc_kg_ci = (toc_ci * area_m2)/1000, 
      total_tn_kg = (tn * area_m2)/(1000*1000), # kg
      total_tn_kg_ci = (tn_ci * area_m2)/(1000*1000),
    ) %>% 
    # reformat table
    pivot_longer(-site) %>% 
    mutate(type = ifelse(name %>% str_detect("_ci"), "confint", "value"),
           name = name %>% str_remove("_ci")) %>% 
    pivot_wider(names_from = c(site, type)) %>% 
    relocate(name, vowell_value, vowell_confint) %>% 
    mutate_if(is.numeric, format, digits=3, nsmall = 0) %>% 
    mutate(name = name %>% fct_recode(bil_cells_per_m2 = "cell_count",
                                      cell_area_percent_cov = "cell_area", 
                                      toc_g_m2 = "toc",
                                      tn_mg_m2 = "tn", 
                                      `total_cells_10^9` = "total_cells"))
}

bloom_polygon_stats <- ts_custom_wide %>% 
  get_max_s2_rgnd() %>% 
  estimate_bloom_stats(polygon_area_df = bloom_polygon_area)


bloom_polygon_stats %>% 
  kbl() %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```

### compare w MODIS albedo
```{r}
ts_custom_wide %>% 
  filter(date<"2020-09-01") %>% 
  mutate(site = site %>% fct_relevel("vowell")) %>% 
  drop_na(albedo) %>% 
  ggplot(aes(x = date, y = albedo)) +
  geom_point() +
  facet_grid(rows = vars(site)) +
  geom_smooth(se = FALSE) +
  ggtitle("MODIS snow albedo over bloom polygon")
```
Albedo decline in July of 
vowell: 0.5 - 0.35 ~= 0.15
catamount: 0.55 - 0.45 ~= 0.1

Our regression predicts a decline of:
vowell: 0.05
catamount: 0.035


## glacier polygon

```{r}
# from GEE
glacier_polygon_area <- tribble(
  ~site, ~area_m2,
  "vowell", 6.03739e6,
  "cat", 5.84705e6
)

glacier_polygon_stats <- ts_custom_wide %>% 
  get_max_s2_rgnd() %>% 
  estimate_bloom_stats(polygon_area_df = glacier_polygon_area)


glacier_polygon_stats %>% 
  kbl() %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```




### compare with MODIS snow albedo

```{r}
ts_glims_wide %>% 
  filter(date<"2020-09-01") %>% 
  drop_na(albedo) %>% 
  ggplot(aes(x = date, y = albedo)) +
  geom_point() +
  facet_grid(rows = vars(site)) +
  geom_smooth(se = FALSE) +
  ggtitle("MODIS snow albedo over glacier polygon")
```


# summer radiative forcing

```{r}
summer_rf <- function(srad_df, delta_albedo_df){
  srad_df %>% 
    filter(month(date) %in% c(7,8)) %>% 
    select(date, srad) %>% 
    left_join(delta_albedo_df, by = "date") %>% 
    mutate(rf = srad * delta_albedo,
           rf_ci = srad * delta_albedo_ci) %>% 
    group_by(site) %>% 
    summarise(rf_wm2 = mean(rf), rf_ci = mean(rf_ci))
}
```


Using S2 mean RGND within bloom polygon

```{r}
s2_interp_d_albedo <- s2_interp %>% 
  mutate(delta_albedo = 0.9 * rgnd, 
         delta_albedo_ci = 0.34 * rgnd) %>% 
  select(-albedo)

wx_clean %>% 
  select(site, date, srad) %>% 
  summer_rf(delta_albedo_df = s2_interp_d_albedo)
```

Mean July-August RF:

vowell: 10.6 +/- 4 W/m2 
cat: 7.4 +/- 2.8 W/m2






# compare RGND time series with different polygons


```{r}
plot_dat2 <- ts_custom_wide %>% add_column(polygon = "bloom") %>% 
  bind_rows(ts_modis_wide %>% add_column(polygon = "modis")) %>% 
  bind_rows(ts_glims_wide %>% add_column(polygon = "glacier")) %>% 
  filter(platform %in% c("s2", "l8", "terra")) %>% 
  mutate(platform = platform %>% 
           fct_relevel(c("s2", "l8", "terra")) %>%# order factors
           fct_recode("Sentinel-2"="s2",
                      "Landast-8"="l8",
                      "MODIS/Terra"="terra")) 


plot_dat2 %>% 
  ggplot(aes(x = date, y = rgnd, color = polygon, shape = polygon)) +
  facet_grid(rows = vars(site %>% fct_relevel("vowell")), cols = vars(platform)) +
  geom_point(alpha = 0.5) +
  # geom_line(aes(group=date)) +
  labs(x = "Date", y = "RGND") +
  theme_minimal() +
  geom_hline(yintercept = 0)
```







# RGND histogram (s2 density plot)
```{r}
grid_cell_count <- hist_clean %>% 
  group_by(site) %>% 
  summarise(grid_cell_count = sum(rgnd_count, na.rm = TRUE))
grid_cell_count

hist_clean %>% 
  drop_na(rgnd_count) %>% # remove empty bins
  left_join(grid_cell_count) %>% 
  mutate(percent_algae_area = (rgnd_count / grid_cell_count) *100) %>% 
  ggplot(aes(x = band_value, y = percent_algae_area)) +
  geom_line() +
  facet_grid(rows = vars(site)) +
  labs(y = "% algae area", x = "RGND")
```
```{r}
# reverse engineer the raw obs, hacky
raw_vals_vec_cat <- hist_clean %>% 
  filter(site == "cat") %>% 
  drop_na(rgnd_count) %>% 
  rowwise() %>% 
  mutate(reps = list(rep(band_value, rgnd_count))) %>% # ->x;x[9,4] %>% pull() # sanity check
  select(reps) %>% 
  unlist() # as list

raw_vals_vec_vowell <- hist_clean %>% 
  filter(site == "vowell") %>% 
  drop_na(rgnd_count) %>% 
  rowwise() %>% 
  mutate(reps = list(rep(band_value, rgnd_count))) %>% # ->x;x[9,4] %>% pull() # sanity check
  select(reps) %>% 
  unlist() # as list

raw_vals_tbl <- tibble(rgnd = c(raw_vals_vec_cat, raw_vals_vec_vowell),
                       site = c(rep("cat", length(raw_vals_vec_cat)), 
                                rep("vowell", length(raw_vals_vec_vowell))) )

raw_vals_tbl %>% 
  group_by(site) %>% 
  summarise(mean = mean(rgnd), max = max(rgnd), min = min(rgnd), sd = sd(rgnd)) %>% 
  mutate(mean_cell_percent = mean * 154, 
         mean_toc = 12.9*mean,
         mean_cell = mean * 5.4*mean)

# # santiy check: compute mean manually from the histogram data
# hist_clean %>% 
#   drop_na(rgnd_count) %>% 
#   left_join(grid_cell_count) %>% 
#   group_by(site) %>% 
#   mutate(weighted = band_value*rgnd_count) %>% 
#   summarise(mean = sum(weighted)/grid_cell_count) %>% 
#   distinct() # cat is 0.034, vowell is 0.06

```

~99% (2 SD) of values from 0 to 0.12 (vowell), and -0.005 to 0.075
0.06-0.03*2
0.035-0.02


During our Catamount transect we observed small patches of concentrated red snow ~10-100 cm wide interspersed on white snow (Supp), but this patchiness was less apparent in Sentinel-2 images due to 10 m spatial resolution


# weather/albedo metadata


```{r}
my_wx <- ts_custom_wide %>% 
  filter(platform %in% c("terra", "terraSnow")) %>% 
  select(site, date, rgnd, albedo) %>% 
  pivot_longer(-site:-date, values_drop_na =TRUE) %>% 
  pivot_wider() %>% 
  full_join(wx_clean) %>% 
  # mean of both glaciers
  group_by(date) %>%
  select(-site) %>%
  summarise(across(everything(), list(mean = mean), .names = "{.col}")) %>% 
  pivot_longer(-date) %>%
  mutate(name = name %>% fct_relevel("rgnd", "albedo", "tmax", "tmin", "prcp", "srad"))


my_wx %>% 
  ggplot(aes(x = date, y = value)) +
  facet_grid(rows = vars(name), scales = "free") +
  geom_point(alpha = 0.6, data = my_wx %>% filter(name %in% c("rgnd", "albedo"))) +
  theme_minimal() +
  geom_smooth(color = "red", 
              data = my_wx %>% 
                filter(name =="rgnd") %>% 
                # set Jul 1 to 0
                bind_rows(tibble(date = ymd("2020-07-01", "2020-07-02", "2020-07-03", "2020-07-04", "2020-07-05"), 
                                 name = "rgnd", value=0)) %>% 
                mutate(name = name %>% fct_relevel("rgnd"))) +
  geom_smooth(data = my_wx %>% filter(name =="albedo"), color = "black") +
  geom_line(alpha = 0.5, data = my_wx %>% filter(!(name %in% c("rgnd", "albedo")))) +
  # add horiz line to some facets
  geom_hline(aes(yintercept = 0), data = my_wx %>% filter(!(name %in% c("albedo", "srad"))) ) 
```



# Applying G-D melt model, algae melt = 0.13 sqrt(area in cm2/l)

```{r}
melt <- function(rgnd){
  cell_dens <- rgnd * 1234 # my model, in cm2/L
  0.13*sqrt(cell_dens)
}
melt(0.016) # vowell
melt(0.021) # cat

```
```{r}
melt(0.05)# 10.2 mm SWE/day
melt(0.039) #9 mm SWE/day
```
1 cm water over the entire bloom polygon area, is a volume of 
```{r}
# in units of m2 and m
1.6e6*0.012 # 19,000 m3/day
2.1e6*0.009 # 19,000 m3 per day
```

