---
title: "Satellite image comparison"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
library(here)
library(fs)
library(janitor)
library(lubridate)
library(tsibble)
library(splines)
library(zoo)
library(kableExtra)
library(GGally)
library(ggpubr)
library(broom)
```


Not the most elegant solution to download each time series into a seperate csv, but working with what I've got. 

```{r}
# modis polygon
ts_modis <- dir_ls(here("data/google_earth_engine/modis_grid_cells"), regexp = ".csv$") %>% 
  map_df(read_csv, id = "filename", col_types = cols())

# bloom polygon
ts_custom <- dir_ls(here("data/google_earth_engine/custom_polygon"), regexp = ".csv$") %>% 
  map_df(read_csv, id = "filename", col_types = cols())

# bloom polygon pix count
ts_custom_count <- dir_ls(here("data/google_earth_engine/custom_polygon/count"), regexp = "count.csv$") %>% 
  map_df(read_csv, id = "filename", col_types = cols())

# glacier polygon
ts_glims <- dir_ls(here("data/google_earth_engine/glimsGeom"), regexp = ".csv$") %>% 
  map_df(read_csv, id = "filename", col_types = cols())

# histogram of RGND during peak bloom week
hist <- dir_ls(here("data/google_earth_engine/hist/"), regexp = "*hist.csv$") %>% 
  map_df(read_csv, id = "filename", col_types = cols())

# wx
wx <- dir_ls(here("data/google_earth_engine/wx"), regexp = "*wx.csv$") %>% 
  map_df(read_csv, id = "filename", col_types = cols())
```

Coefs for biomass and albedo regression models
```{r}
biomass_coefs <- read_csv(here("data/abundance_coefs.csv"))
albedo_coefs <- read_csv(here("data/albedo_coefs.csv")) %>% 
  add_column("response"= "albedo") # add column to match biomass_coefs

```


# tidy

```{r}
# parse site name, date
my_tidy <- function(df){
  df %>% 
    clean_names() %>% 
    mutate(site = filename %>% basename() %>% str_remove(".csv") %>% str_split_i("_",1),
           date = mdy(system_time_start))
}

# parse platform name, apply cloudmask
ts_tidy <- function(df){
  df %>% 
     mutate(platform = basename(filename) %>% 
              str_split_i("_", 2) %>% 
              str_remove(".csv"),
         rgnd = if_else(is.na(rgnd), rgnd_cloudmasked, rgnd)) %>% 
  select(platform, site, date, rgnd, albedo)
}


ts_modis_wide <- ts_modis %>% 
  my_tidy() %>% 
  ts_tidy()

ts_custom_wide_unfilt <- ts_custom %>% 
  my_tidy() %>% 
  ts_tidy()

ts_custom_count_wide <- ts_custom_count %>% 
  my_tidy() %>% 
  mutate(platform = basename(filename) %>% 
              str_split_i("_", 2) %>% 
              str_remove(".csv"), 
         date = mdy(system_time_start),
         .keep = "unused") %>% 
  rename(count  = rgnd_cloudmasked) %>% 
  select(platform, site, date, count)
  

ts_glims_wide <- ts_glims %>% 
  my_tidy() %>% 
  ts_tidy()

hist_clean <- hist %>% 
  clean_names() %>% 
  mutate(site = basename(filename) %>% str_split_i("_", 1), .keep = "unused")

wx_clean <- wx %>% 
  my_tidy() %>% 
  select(-filename, -system_time_start) %>% 
  pivot_longer(-site:-date, values_drop_na =TRUE) %>% 
  pivot_wider()
  
```


# n pix per image

```{r}
ts_custom_wide_unfilt %>% 
  inner_join(ts_custom_count_wide) %>% 
  ggplot(aes(count)) +
  geom_histogram() +
  facet_wrap(vars(platform), scales = "free")
```

```{r}
ppdat <- ts_custom_wide_unfilt %>% 
  inner_join(ts_custom_count_wide) %>% 
  mutate(platform = platform %>% fct_recode("MODIS/Aqua"="aqua","Sentinel-2"="s2", "Landsat-8"="l8", "MODIS/Terra"="terra") %>% fct_relevel("Sentinel-2", "Landsat-8", "MODIS/Terra"),
         site = site %>% fct_relevel("vowell") %>% fct_recode("Catamount"="cat", "Vowell"="vowell"))

ppdat %>% 
  ggplot(aes(y = count, x = date, color = site)) +
  geom_point(alpha = 0.4) +
  geom_line()+
  facet_wrap(vars(platform), scales = "free_y") +
  scale_colour_brewer(palette = "Dark2")+
  labs(y = "Valid pixel count", x = "Date", tag="A") +
  geom_hline(aes(yintercept = 3), data = ppdat %>% filter(platform %>% str_detect("MODIS"))) +
  geom_hline(aes(yintercept = 15000), data = ppdat %>% filter(platform =="Sentinel-2")) +
  geom_hline(aes(yintercept = 800), data = ppdat %>% filter(platform =="Landsat-8")) 
```


```{r}
count(ts_custom_wide_unfilt, platform)

modis_thresh <- 3
s2_thresh <- 15000
l8_thresh <- 800

ts_custom_wide <-
  ts_custom_wide_unfilt %>% 
  left_join(ts_custom_count_wide) %>% 
  left_join(tribble(
    ~platform, ~thresh,
    "l8",l8_thresh,
    "s2",s2_thresh,
    "terra",modis_thresh,
    "aqua", modis_thresh
  )) %>% #count(platform) # check before
  # hack to filter by thresh without discarding the time series that don't have count/thresh data
  filter(if_else(!is.na(count), count>=thresh, TRUE)) #%>% count(platform) # check after, only should remove S2/L8/Terra
nrow(ts_custom_wide)
count(ts_custom_wide, platform)
```



# n obs per platform

```{r}
n_img_kable <- ts_custom_wide %>% 
  group_by(site, platform) %>% 
  summarise(n_images = n()) %>% 
  ungroup() %>% 
  filter(platform!="terraSnow", platform!="aquaSnow") %>% 
  mutate(platform = platform %>% fct_recode("Landsat-8"="l8",
                                            "Sentinel-2"="s2",
                                            "MODIS/Terra"="terra",
                                            "MODIS/Aqua"="aqua")) %>% 
  pivot_wider(names_from = site, values_from = n_images) %>% 
  kbl(caption = "B. Number of cloud-free satellite images per site used in analysis", col.names = c("Platform", "Catamount", "Vowell")) %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```
```{r}
n_img_kable %>% 
  save_kable(here("figs/n_images_table.pdf"))
```
```{bash}
# the density parameter must come before the input file
convert \
  -density 300 \
  ../figs/n_images_table.pdf \
  -sharpen 0x1.0 \
  -quality 100 \
  ../figs/n_images_table.png
```
N images July 5 - Sept 1
```{r}
ts_custom_wide %>% 
  filter(date>"2020-07-05", date<"2020-09-01") %>% 
  group_by(site, platform) %>% 
  summarise(n_images = n()) %>% 
  ungroup() %>% 
  filter(platform=="s2")

```

# fig6 plot


```{r}
# make pretty for plotting
plot_dat_4 <- ts_custom_wide %>% 
  filter(platform %in% c("s2", "l8", "terra", "aqua"),
         # remove sept 8 for catamount, looks like snow buried the algae on this date. Our index not calibrated for lightly buried snow algae
         !(platform=="s2" & site=="cat" & date==ymd("2020-09-08")), 
         # full polygon was set as "not snow" and therefore set RGND=0
         !(platform=="aqua" & site == "vowell" & date==ymd("2020-07-22")), 
         # all but one pixels set as not snow, only 1 valid RGND pixel, 
         ) %>% 
  mutate(platform = platform %>% 
           fct_recode("Sentinel-2"="s2",
                      "Landsat-8"="l8",
                      "MODIS/Terra"="terra",
                      "MODIS/Aqua"="aqua") %>% 
           fct_relevel(c("Sentinel-2", "Landsat-8", "MODIS/Terra")),
         site = site %>% fct_relevel("vowell")) %>% 
  mutate(site = fct_recode(site, "Vowell"="vowell","Catamount"="cat")) %>% 
  dplyr::rename(Platform = platform)
plot_dat_4$Platform %>% levels()
plot_dat <- plot_dat_4 %>% filter(Platform!="MODIS/Aqua")

gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

p1 <- plot_dat %>% 
  ggplot(aes(x = date, y = rgnd, color = Platform, shape = Platform)) +
  facet_grid(rows = vars(site)) +
  geom_point(alpha =0.7) +
  labs(x = "Date", y = "RGND") +
  theme_minimal() +
  geom_hline(yintercept = 0) +
  theme(panel.spacing = unit(2, "lines")) 
p1
```


## sentinel linear interpolation

```{r}
s2_interp <- plot_dat %>% 
  filter(Platform=="Sentinel-2", 
         # just July-Aug data
         date >= "2020-07-01",
         date <= "2020-09-12") %>% 
  # set 0 on Jul 1 and Sept 1
  bind_rows(tribble(
    ~site, ~date, ~rgnd,
    "Catamount", ymd("2020-07-05"), 0,
    "Catamount", ymd("2020-09-01"), 0,
    "Vowell", ymd("2020-07-05"), 0,
    "Vowell", ymd("2020-09-01"), 0
  )) %>% 
  # convert to tsibble to use handy fill_gaps()
  as_tsibble(key = site) %>%
  # make missing values explicit NA
  fill_gaps() %>% 
  # linear interpolate NA
  mutate(rgnd = zoo::na.approx(rgnd)) %>% 
  as_tibble() %>% 
  # add stuff back in for plotting
  mutate(Platform="Sentinel-2",
         site = site %>% fct_relevel("Vowell"))
```

## export final plot

```{r}
p1 +
  # geom_line(stat="smooth", method = "loess", data= curve_dat, alpha = 0.5, size = 1, color = "purple")
  geom_line(data= s2_interp, alpha = 0.5, size = 1, color = gg_color_hue(3)[1]) +
  # re plot points, so the points are on top of the line
  geom_point(alpha = 0.7)

```

```{r}
ggsave(here("figs/figs_v2/7_rgnd_ts/satellite_time_series.pdf"), height = 5, width = 7)
```




# Compare satellite RGND w/in 5 day window

## pairs plot
```{r}
date_bins <- plot_dat_4 %>% 
  # bin time into discrete 4 day intervals
  mutate(date_bin = date %>% yday() %>% cut_width(width=5)) %>% 
  group_by(Platform, site, date_bin) %>% 
  summarise(rgnd = mean(rgnd)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = "Platform", values_from = "rgnd")

# check that cut_width works as expected
date_bins %>% 
  distinct(date_bin) %>% 
  arrange(date_bin)
```

Comparison of mean values for each 5 day time interval
```{r}
my_fn <- function(data, mapping, ...){
  ggplot(data = data, mapping = mapping) + 
    geom_point(aes(color = site)) + 
    geom_line(stat="smooth", method="lm", ...) +
    scale_x_continuous(limits = c(0,0.04))+
    scale_y_continuous(limits = c(0,0.04))+
    scale_colour_brewer(palette = "Dark2")+
    geom_abline(slope = 1, intercept = 0, linetype="dashed", color = "grey50") 
}

date_bins %>%
  ggpairs(columns = 3:6,
          lower = list(continuous = my_fn))
```

```{r}
mod_s2_ratio <- date_bins %>% 
  clean_names() %>% 
  rowwise() %>% # a special form of grouping
  mutate(mean_modis = mean(c(modis_terra, modis_aqua), na.rm=T),
         mod_s2_ratio = mean_modis/sentinel_2) %>% 
  ungroup()
mod_s2_ratio

mod_s2_ratio %>% 
  filter(!(is.na(mod_s2_ratio)), !(is.infinite(mod_s2_ratio))) %>% 
  summarise(mean(mod_s2_ratio), sd(mod_s2_ratio))
  
```



```{r}
# ppa <- date_bins %>%
#   ggplot(aes(y = `Sentinel-2`, x = `MODIS/Terra`)) +
#   geom_point(aes(color = site), size = 2.5) +
#   lims(x = c(0, 0.04), y = c(0, 0.04)) +
#   geom_smooth(method = "lm", se=FALSE) +
#   theme_minimal() +
#   labs(y = "Sentinel-2 RGND", x = "MODIS/Terra RGND", tag = "A") +
#   geom_abline(slope = 1, intercept = 0, linetype="dashed", color = "grey50") +
#   scale_colour_brewer(palette = "Dark2")
# ppa  
# 
# ppb <- date_bins %>%
#   ggplot(aes(y = `Sentinel-2`, x = `Landsat-8`)) +
#   geom_point(aes(color = site), size = 2.5) +
#   lims(x = c(0, 0.04), y = c(0, 0.04)) +
#   geom_smooth(method = "lm", se=FALSE) +
#   theme_minimal() +
#   labs(y = "Sentinel-2 RGND", x = "Landsat-8", tag = "B") +
#   geom_abline(slope = 1, intercept = 0, linetype="dashed", color = "grey50") +
#   scale_colour_brewer(palette = "Dark2")
# ppb
# 
# ppc <- date_bins %>%
#   ggplot(aes(y = `Landsat-8`, x = `MODIS/Terra`)) +
#   geom_point(aes(color = site), size = 2.5) +
#   lims(x = c(0, 0.04), y = c(0, 0.04)) +
#   geom_smooth(method = "lm", se=FALSE) +
#   theme_minimal() +
#   labs(y = "Landsat-8 RGND", x = "MODIS/Terra RGND", tag = "C") +
#   geom_abline(slope = 1, intercept = 0, linetype="dashed", color = "grey50") +
#   scale_colour_brewer(palette = "Dark2")
# ppc
# 
# legend <- cowplot::get_legend(ppa)
# ggarrange(ppa + theme(legend.position = "none"), ppb + theme(legend.position = "none"), ppc + theme(legend.position = "none"), legend)
```
```{r}
# ggsave(here("figs/rgnd_compare.png"))
```


## peak bloom boxplot


```{r}
peak_yday <- yday(ymd('2020-07-28'))
# i can see this date bin is (208,212], thats Jul 27 -Jul 30


plot_dat %>% 
  filter(date>ymd("2020-07-26"), date <= ymd("2020-07-30")) %>% 
  ggplot(aes(x = Platform, y = rgnd)) +
  geom_boxplot() +
  facet_wrap(vars(site))
```
defining "peak week" as Jul 27 -Jul 30

# algae stats during peak week


## bloom polygon

Apply regression estimates:


```{r}
# from GEE
bloom_polygon_area <- tribble(
  ~site, ~area_m2,
  "vowell", 1684302,
  "cat", 1918716
)
```


```{r}
get_max_s2_rgnd <- function(df){
  df %>% 
  filter(platform == "s2") %>% 
  group_by(site) %>% 
  summarise(rgnd = max(rgnd, na.rm = TRUE))
}
ts_custom_wide %>% 
  filter(platform == "s2") %>% 
  arrange(-rgnd)

# helper function to extract estimates
get_coef <- function(coef_tbl, rr, col){
  coef_tbl %>% 
    filter(term != "(Intercept)",
           response == rr) %>%
    pull({{col}})
}

estimate_bloom_stats <- function(rgnd_df, polygon_area_df){
  rgnd_df %>%
    left_join(polygon_area_df) %>% 
    mutate(
      cell_count = get_coef(biomass_coefs, "cell abundance", estimate) * rgnd^2,
      cell_count_se = get_coef(biomass_coefs, "cell abundance", std.error) * rgnd^2,
      cell_area = get_coef(biomass_coefs, "cell area", estimate) * rgnd^2, 
      cell_area_se = get_coef(biomass_coefs, "cell area", std.error) * rgnd^2,
      toc = get_coef(biomass_coefs, "toc", estimate) * rgnd, # g/m2
      toc_se = get_coef(biomass_coefs, "toc", std.error) * rgnd,
      tn = get_coef(biomass_coefs, "tn", estimate) * rgnd, # mg/m2
      tn_se = get_coef(biomass_coefs, "tn", std.error) * rgnd,
      delta_albedo = get_coef(albedo_coefs, "albedo", estimate) * rgnd, # delta albedo
      delta_albedo_se = get_coef(albedo_coefs, "albedo", std.error) * rgnd, 
      total_cells = cell_count * area_m2, # bil cells
      total_cells_se = cell_count_se * area_m2, 
      total_toc_kg = (toc * area_m2)/1000, # kg
      total_toc_kg_se = (toc_se * area_m2)/1000, 
      total_tn_kg = (tn * area_m2)/(1000*1000), # kg
      total_tn_kg_se = (tn_se * area_m2)/(1000*1000),
    ) %>% 
    # reformat table
    pivot_longer(-site) %>% 
    mutate(type = ifelse(name %>% str_detect("_se"), "se", "value"),
           name = name %>% str_remove("_se")) %>% 
    pivot_wider(names_from = c(site, type)) %>% 
    relocate(name, vowell_value, vowell_se) %>% # put vowell before catamount, consistant w rest of manuscript
    mutate_if(is.numeric, format, digits=3, nsmall = 0) %>% 
    mutate(name = name %>% fct_recode(bil_cells_per_m2 = "cell_count",
                                      cell_area_percent_cov = "cell_area", 
                                      toc_g_m2 = "toc",
                                      tn_mg_m2 = "tn", 
                                      `total_cells_10^9` = "total_cells"))
}

bloom_polygon_stats <- ts_custom_wide %>% 
  get_max_s2_rgnd() %>% 
  estimate_bloom_stats(polygon_area_df = bloom_polygon_area)


bloom_kable <- bloom_polygon_stats %>% 
  mutate(name=name %>% fct_recode(
    "RGND"="rgnd",
    "Bloom area (m^2)" = "area_m2",
    "Cell area density (10^9 cells/m^2)" ="bil_cells_per_m2",
    "Cell area % cover" ="cell_area_percent_cov",
    "TOC (g/m^2)" = "toc_g_m2",
    "TN (mg/m^2)" = "tn_mg_m2",
    "Delta albedo" = "delta_albedo",
    "Total cells (10^9)"="total_cells_10^9",
    "Total TOC (kg)" = "total_toc_kg",
    "Total TN (kg)" = "total_tn_kg"
  )) %>% 
  kbl(caption = "A. Bloom polygon averages", col.names = c("", "Vowell", "Vowell SE", "Catamount", "Catamount SE")) %>% 
  kable_classic(full_width = F, html_font = "Cambria")
bloom_kable
```

Compute mean values

```{r}
bloom_polygon_stats %>% 
  mutate(across(vowell_value:cat_se, as.numeric)) %>% 
  pivot_longer(-name, names_to = "s", values_drop_na = TRUE) %>% 
  separate(s, into = c("site", "stat")) %>% 
  group_by(name, stat) %>% 
  summarise(mean = mean(value)) %>% 
  pivot_wider(names_from= "stat", values_from = "mean")

```


```{r}
bloom_kable %>% 
  save_kable(here("figs/bloom_poly_stats_table.pdf"))
```
```{bash}
# the density parameter must come before the input file
convert \
  -density 300 \
  ../figs/bloom_poly_stats_table.pdf \
  -sharpen 0x1.0 \
  -quality 100 \
  ../figs/bloom_poly_stats_table.png
```

### compare w MODIS albedo
```{r}
ts_custom_wide %>% 
  filter(date<"2020-09-01") %>% 
  mutate(site = site %>% fct_relevel("vowell")) %>% 
  drop_na(albedo) %>% 
  ggplot(aes(x = date, y = albedo)) +
  geom_point() +
  facet_grid(rows = vars(site)) +
  geom_smooth(se = FALSE) +
  ggtitle("MODIS snow albedo over bloom polygon")
```
Extract LOESS estimates
```{r}
vowell_dat <- ts_custom_wide %>% filter(platform =="terraSnow", site=="vowell")
cat_dat <- ts_custom_wide %>% filter(platform =="terraSnow", site=="cat")

dates_in_summer <- ts_custom_wide %>% distinct(date)
ydays_in_summer <- dates_in_summer %>% mutate(date = yday(date)) %>% pull()

vowell_mod <- loess(albedo ~ yday(date), data = vowell_dat)
cat_mod <- loess(albedo ~ yday(date), data = cat_dat)

vowell_aug <- vowell_mod %>% broom::augment(vowell_dat)
cat_aug <- cat_mod %>% augment(cat_dat)

cat_aug %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y=.fitted)) +
  geom_point(aes(y = albedo))
```
```{r}
vowell_grid <- yday(ymd("2020-07-05")):yday(ymd("2020-07-18"))
cat_grid <- yday(ymd("2020-07-05")):yday(ymd("2020-07-30"))

vowell_pred <- vowell_mod %>% predict(vowell_grid)
cat_pred <- cat_mod %>% predict(cat_grid)  

vowell_diff <- vowell_pred[1] - vowell_pred[vowell_pred==min(vowell_pred)]
cat_diff <- cat_pred[1] - cat_pred[cat_pred==min(cat_pred)]
vowell_diff
cat_diff

```




## glacier polygon

```{r}
# from GEE
glacier_polygon_area <- tribble(
  ~site, ~area_m2,
  "vowell", 6.03739e6,
  "cat", 5.84705e6
)

glacier_polygon_stats <- ts_glims_wide %>% 
  get_max_s2_rgnd() %>% 
  estimate_bloom_stats(polygon_area_df = glacier_polygon_area)


glims_kable <- glacier_polygon_stats %>% mutate(name=name %>% fct_recode(
    "RGND"="rgnd",
    "Glacier area (m^2)" = "area_m2",
    "Cell area density (10^9 cells/m^2)" ="bil_cells_per_m2",
    "Cell area % cover" ="cell_area_percent_cov",
    "TOC (g/m^2)" = "toc_g_m2",
    "TN (mg/m^2)" = "tn_mg_m2",
    "Delta albedo" = "delta_albedo",
    "Total cells (10^9)"="total_cells_10^9",
    "Total TOC (kg)" = "total_toc_kg",
    "Total TN (kg)" = "total_tn_kg"
  )) %>% 
  kbl(caption = "B. Glacier polygon averages", col.names = c("", "Vowell", "Vowell SE", "Catamount", "Catamount SE")) %>% 
  kable_classic(full_width = F, html_font = "Cambria")
glims_kable
```



```{r}
glims_kable %>% 
  save_kable(here("figs/glims_poly_stats_table.pdf"))
```
```{bash}
# the density parameter must come before the input file
convert \
  -density 300 \
  ../figs/glims_poly_stats_table.pdf \
  -sharpen 0x1.0 \
  -quality 100 \
  ../figs/glims_poly_stats_table.png
```

### compare with MODIS snow albedo

```{r}
ts_glims_wide %>% 
  filter(date<"2020-09-01") %>% 
  drop_na(albedo) %>% 
  ggplot(aes(x = date, y = albedo)) +
  geom_point() +
  facet_grid(rows = vars(site)) +
  geom_smooth(se = FALSE) +
  ggtitle("MODIS snow albedo over glacier polygon")
```


# summer radiative forcing

```{r}
summer_rf <- function(srad_df, delta_albedo_df){
  srad_df %>% 
    filter(month(date) %in% c(7,8)) %>% 
    select(date, srad) %>% 
    left_join(delta_albedo_df, by = "date") %>% 
    mutate(rf = srad * delta_albedo,
           rf_se = srad * delta_albedo_se) %>% 
    group_by(site) %>% 
    summarise(rf_wm2 = mean(rf), rf_se = mean(rf_se))
}
```


Using S2 mean RGND within bloom polygon to estimate July-August RF

```{r}
# using the interpolated data with assumed start and end points
s2_interp_d_albedo <- s2_interp %>% 
  mutate( delta_albedo = get_coef(albedo_coefs, "albedo", estimate) * rgnd, 
          delta_albedo_se = get_coef(albedo_coefs, "albedo", std.error) * rgnd) %>% 
  select(-albedo)

wx_clean %>% 
  select(site, date, srad) %>% 
  summer_rf(delta_albedo_df = s2_interp_d_albedo) %>% 
  drop_na()
```

Compute the total energy absorbed by the algae during this time period from DRF

```{r}
energy <- wx_clean %>% 
  filter(date > "2020-07-05", date < "2020-09-01") %>% 
  select(date, srad, dayl) %>% 
  left_join(s2_interp_d_albedo %>% select(date, site, delta_albedo, delta_albedo_se), by = "date") %>% 
  mutate(drf = -1* (srad * delta_albedo), # RF should be positive
         drf_se = srad * delta_albedo_se,
         j_per_m2 = drf * dayl, # since energy = power (RF) * time (s), and SRF is averaged over a 12 hour day
        j_per_m2_se = drf_se * dayl) 


energy %>% 
  ggplot(aes(x = date, y = j_per_m2)) +
  geom_ribbon(aes(ymin = j_per_m2 - j_per_m2_se,
                  ymax = j_per_m2 + j_per_m2_se, group = site, fill = site), 
              alpha = 0.2, show.legend=FALSE) +
  geom_line(aes(color=site)) +
  theme_minimal() +
  labs(x = "Date", y = "Energy absorbed by snow algae albedo (MJ/m2)", color = "Site") +
  scale_colour_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels=function(x)x/1000000)
  

```

```{r}
snow <- read_csv(here("data/snow_stations.csv"), skip = 5)
names(snow) <- c("date", "2D09_SD", "2D09_SW", "2C14P_SD", "2C14P_SW", "2A34P_SD", "2A34P_SW")
snow <- snow %>% mutate(date = date %>% str_sub(1,10) %>% ymd())

snow_long <- snow %>% 
  pivot_longer(-date, values_drop_na = TRUE) %>% 
  separate(name, c("site", "name")) %>% 
  group_by(date, site, name) %>% 
  summarise(value=mean(value)) %>% 
  ungroup()

snow_long %>%
  mutate(year = year(date)) %>% 
  group_by(site) %>% 
  summarise(min(year), max(year))

snow_long %>%
  pivot_wider() %>% 
  mutate(percent = (SW/SD)*100, month = as.factor(month(date))) %>% 
  filter(month %in% 1:5) %>% 
  ggplot(aes(x = month, y  = percent)) +
  geom_boxplot(aes(group = month)) +
  facet_wrap(vars(site)) +
  labs(y = "Percent water")
  
```




The total energy delivered:
```{r}
net_energy <- energy %>% 
  group_by(site) %>% 
  summarise(sum_algal_energy = sum(j_per_m2), sum_algal_se = sum(j_per_m2_se), sum_srad = sum((srad*dayl)))

net_energy

latent_heat_snow <- 3.3*10^5 #J/kg
snow_density <- 500 #kg/m3 

net_energy %>% 
  pivot_longer(-site) %>% 
  mutate(snowmelt_eq_kg = value / latent_heat_snow,# in kg water / m2
         snowdepth_eq_m = snowmelt_eq_kg/snow_density)
```

Convert this per total area?


Computing this another way, taking the averages first and then multiplying srad by delta albedo
```{r}
# mean summer incoming radiation
mean_summer_srad <- wx_clean %>% 
  filter(month(date) %in% c(7,8)) %>% 
  group_by(site) %>% 
  summarise(mean(srad)) %>% 
  mutate(site = site %>% fct_recode("Vowell"="vowell", "Catamount"="cat")) 


# mean summer d albedo
s2_interp_d_albedo %>% 
  group_by(site) %>% 
  summarise(mean(delta_albedo), mean(delta_albedo_se)) %>% 
  left_join(mean_summer_srad) %>% 
  mutate(srf = `mean(delta_albedo)` * `mean(srad)`,
         srf_se = `mean(delta_albedo_se)` * `mean(srad)`)


```




# compare RGND time series with different polygons


```{r, fig.height = 5, fig.width=8}
plot_dat2 <- ts_custom_wide %>% add_column(polygon = "bloom") %>% 
  bind_rows(ts_modis_wide %>% add_column(polygon = "modis")) %>% 
  bind_rows(ts_glims_wide %>% add_column(polygon = "glacier")) %>% 
  filter(platform %in% c("s2", "l8", "terra", "aqua"),
         # same filtering criteria as above (duplicated code... not the best approach)
          !(platform=="s2" & site=="cat" & date==ymd("2020-09-08")), 
         # full polygon was set as "not snow" and therefore set RGND=0
         !(platform=="aqua" & site == "vowell" & date==ymd("2020-07-22")), 
         # discard nonsensical obs-- RGND spike on Sep 30 cannot be due to snow algae, b/c there is no snow.
         !(platform=="aqua" & site == "vowell" & date==ymd("2020-09-30")), 
         # discard nonsensical high RGND in early June for aqua, only one pixel, no algae in S2 and L8 this early
         !(platform=="aqua" & site == "vowell" & date<ymd("2020-07-01"))
         ) %>% 
  mutate(platform = platform %>% 
           fct_recode("Sentinel-2"="s2",
                      "Landsat-8"="l8",
                      "MODIS/Terra"="terra",
                      "MODIS/Aqua"="aqua")) %>% 
  mutate(site = site %>% fct_recode("Catamount"="cat", "Vowell"="vowell")) %>% 
  # add data for illustrative purposes, set to 0 on Jun 1 where data missing
  bind_rows(tribble(
    ~platform, ~site, ~date, ~rgnd, ~albedo, ~count, ~thresh, ~polygon,
    "Sentinel-2", "Catamount", ymd("2020-07-05"), 0,NA,NA,NA,"bloom",
    "Sentinel-2", "Catamount", ymd("2020-07-05"), 0,NA,NA,NA,"modis",
    "Sentinel-2", "Catamount", ymd("2020-07-05"), 0,NA,NA,NA,"glacier",
    "Sentinel-2", "Vowell", ymd("2020-07-05"), 0,NA,NA,NA,"bloom",
    "Sentinel-2", "Vowell", ymd("2020-07-05"), 0,NA,NA,NA,"modis",
    "Sentinel-2", "Vowell", ymd("2020-07-05"), 0,NA,NA,NA,"glacier",
    "Landsat-8", "Catamount", ymd("2020-07-05"), 0,NA,NA,NA,"bloom",
    "Landsat-8", "Catamount", ymd("2020-07-05"), 0,NA,NA,NA,"modis",
    "Landsat-8", "Catamount", ymd("2020-07-05"), 0,NA,NA,NA,"glacier",
    "Landsat-8", "Vowell", ymd("2020-07-05"), 0,NA,NA,NA,"bloom",
    "Landsat-8", "Vowell", ymd("2020-07-05"), 0,NA,NA,NA,"modis",
    "Landsat-8", "Vowell", ymd("2020-07-05"), 0,NA,NA,NA,"glacier",
    "MODIS/Terra", "Catamount", ymd("2020-07-05"), 0,NA,NA,NA,"bloom",
    "MODIS/Terra", "Catamount", ymd("2020-07-05"), 0,NA,NA,NA,"modis",
    "MODIS/Terra", "Catamount", ymd("2020-07-05"), 0,NA,NA,NA,"glacier",
    "MODIS/Terra", "Vowell", ymd("2020-07-05"), 0,NA,NA,NA,"bloom",
    "MODIS/Terra", "Vowell", ymd("2020-07-05"), 0,NA,NA,NA,"modis",
    "MODIS/Terra", "Vowell", ymd("2020-07-05"), 0,NA,NA,NA,"glacier",
    "MODIS/Aqua", "Catamount", ymd("2020-07-05"), 0,NA,NA,NA,"bloom",
    "MODIS/Aqua", "Catamount", ymd("2020-07-05"), 0,NA,NA,NA,"modis",
    "MODIS/Aqua", "Catamount", ymd("2020-07-05"), 0,NA,NA,NA,"glacier",
    "MODIS/Aqua", "Vowell", ymd("2020-07-05"), 0,NA,NA,NA,"bloom",
    "MODIS/Aqua", "Vowell", ymd("2020-07-05"), 0,NA,NA,NA,"modis",
    "MODIS/Aqua", "Vowell", ymd("2020-07-05"), 0,NA,NA,NA,"glacier",
    
    "Sentinel-2", "Catamount", ymd("2020-09-01"), 0,NA,NA,NA,"bloom",
    "Sentinel-2", "Catamount", ymd("2020-09-01"), 0,NA,NA,NA,"modis",
    "Sentinel-2", "Catamount", ymd("2020-09-01"), 0,NA,NA,NA,"glacier",
    "Sentinel-2", "Vowell", ymd("2020-09-01"), 0,NA,NA,NA,"bloom",
    "Sentinel-2", "Vowell", ymd("2020-09-01"), 0,NA,NA,NA,"modis",
    "Sentinel-2", "Vowell", ymd("2020-09-01"), 0,NA,NA,NA,"glacier",
    "Landsat-8", "Catamount", ymd("2020-09-01"), 0,NA,NA,NA,"bloom",
    "Landsat-8", "Catamount", ymd("2020-09-01"), 0,NA,NA,NA,"modis",
    "Landsat-8", "Catamount", ymd("2020-09-01"), 0,NA,NA,NA,"glacier",
    "Landsat-8", "Vowell", ymd("2020-09-01"), 0,NA,NA,NA,"bloom",
    "Landsat-8", "Vowell", ymd("2020-09-01"), 0,NA,NA,NA,"modis",
    "Landsat-8", "Vowell", ymd("2020-09-01"), 0,NA,NA,NA,"glacier",
    "MODIS/Terra", "Catamount", ymd("2020-09-01"), 0,NA,NA,NA,"bloom",
    "MODIS/Terra", "Catamount", ymd("2020-09-01"), 0,NA,NA,NA,"modis",
    "MODIS/Terra", "Catamount", ymd("2020-09-01"), 0,NA,NA,NA,"glacier",
    "MODIS/Terra", "Vowell", ymd("2020-09-01"), 0,NA,NA,NA,"bloom",
    "MODIS/Terra", "Vowell", ymd("2020-09-01"), 0,NA,NA,NA,"modis",
    "MODIS/Terra", "Vowell", ymd("2020-09-01"), 0,NA,NA,NA,"glacier",
    "MODIS/Aqua", "Catamount", ymd("2020-09-01"), 0,NA,NA,NA,"bloom",
    "MODIS/Aqua", "Catamount", ymd("2020-09-01"), 0,NA,NA,NA,"modis",
    "MODIS/Aqua", "Catamount", ymd("2020-09-01"), 0,NA,NA,NA,"glacier",
    "MODIS/Aqua", "Vowell", ymd("2020-09-01"), 0,NA,NA,NA,"bloom",
    "MODIS/Aqua", "Vowell", ymd("2020-09-01"), 0,NA,NA,NA,"modis",
    "MODIS/Aqua", "Vowell", ymd("2020-09-01"), 0,NA,NA,NA,"glacier"
  )) %>% 
  # order factors last, bind_row() resets the factor order
  mutate(platform = platform %>% fct_relevel(c("Sentinel-2", "Landsat-8", "MODIS/Terra", "MODIS/Aqua")))

plot_dat2 %>% 
  ggplot(aes(x = date, y = rgnd, color = polygon)) +
  facet_grid(rows = vars(site %>% fct_relevel("Vowell")), cols = vars(platform)) +
  geom_hline(yintercept = 0) +
  geom_line(stat="smooth", method ="loess", se=FALSE, alpha = 0.7, data = plot_dat2) +#method = "lm", formula = y~poly(x,2)
  # highlight where data was added
  geom_point(color = "#0cc10f", alpha = 0.4, shape = 1, size = 3, data = plot_dat2 %>% 
               filter(date %in% c(ymd("2020-07-05"),ymd("2020-09-01")))) +
  # hack to double plot to increase the thickness of line
  geom_point(color = "#0cc10f", alpha = 0.4, shape = 1, size = 4, data = plot_dat2 %>% 
               filter(date %in% c(ymd("2020-07-05"),ymd("2020-09-01")))) +
  # the actual data
  geom_point(size = 2, alpha=0.3, data = plot_dat2 %>% filter(!(date %in% c(ymd("2020-07-05"),ymd("2020-09-01"))))) +
  # geom_line(aes(group=date)) +
  labs(x = "Date", y = "RGND") +
  theme_minimal() +
  # scale_shape_manual(values=c(21, 24, 22, 23), guide = "none") +
  scale_color_manual(values = c("black", "#0cbec1", "#c10cbe")) +
  ylim(0, 0.04)
```

```{r}
# ggsave(here("figs/polygon_compare.png"), height=5, width=8)
```






# RGND histogram (s2 density plot)

```{r}
hist_clean %>% 
  group_by(site) %>% 
  summarise(max(band_value))
```

```{r}
grid_cell_count <- hist_clean %>% 
  group_by(site) %>% 
  summarise(grid_cell_count = sum(rgnd_count, na.rm = TRUE))
grid_cell_count

hist_clean %>% 
  drop_na(rgnd_count) %>% # remove empty bins
  left_join(grid_cell_count) %>% 
  mutate(percent_algae_area = (rgnd_count / grid_cell_count) *100) %>% 
  ggplot(aes(x = band_value, y = percent_algae_area)) +
  geom_line() +
  facet_grid(rows = vars(site)) +
  labs(y = "% algae area", x = "RGND")
```
```{r}
# reverse engineer the raw obs, hacky
raw_vals_vec_cat <- hist_clean %>% 
  filter(site == "cat") %>% 
  drop_na(rgnd_count) %>% 
  rowwise() %>% 
  mutate(reps = list(rep(band_value, rgnd_count))) %>% # ->x;x[9,4] %>% pull() # sanity check
  select(reps) %>% 
  unlist() # as list

raw_vals_vec_vowell <- hist_clean %>% 
  filter(site == "vowell") %>% 
  drop_na(rgnd_count) %>% 
  rowwise() %>% 
  mutate(reps = list(rep(band_value, rgnd_count))) %>% # ->x;x[9,4] %>% pull() # sanity check
  select(reps) %>% 
  unlist() # as list

raw_vals_tbl <- tibble(rgnd = c(raw_vals_vec_cat, raw_vals_vec_vowell),
                       site = c(rep("cat", length(raw_vals_vec_cat)), 
                                rep("vowell", length(raw_vals_vec_vowell))) )

raw_vals_tbl %>% 
  group_by(site) %>% 
  summarise(mean = mean(rgnd), max = max(rgnd), min = min(rgnd), sd = sd(rgnd)) %>% 
  mutate(mean_cell_percent = mean * 154, 
         mean_toc = 12.9*mean,
         mean_cell = mean * 5.4*mean)

# # santiy check: compute mean manually from the histogram data
# hist_clean %>% 
#   drop_na(rgnd_count) %>% 
#   left_join(grid_cell_count) %>% 
#   group_by(site) %>% 
#   mutate(weighted = band_value*rgnd_count) %>% 
#   summarise(mean = sum(weighted)/grid_cell_count) %>% 
#   distinct() # cat is 0.034, vowell is 0.06

```

~99% (2 SD) of values from 0 to 0.12 (vowell), and -0.005 to 0.075
0.06-0.03*2
0.035-0.02


During our Catamount transect we observed small patches of concentrated red snow ~10-100 cm wide interspersed on white snow (Supp), but this patchiness was less apparent in Sentinel-2 images due to 10 m spatial resolution


```{r}
raw_vals_tbl %>% 
  ggplot(aes(rgnd)) +
  geom_histogram(binwidth = 0.008)+
  facet_grid(rows = vars(site %>% 
                           fct_relevel("vowell") %>% 
                           fct_recode("Vowell"="vowell", "Catamount"="cat")), 
             scales = "free") +
  labs(x = "Count", y = "RGND") +
  theme_minimal()
```


# weather/albedo metadata



```{r, fig.height = 10, fig.width=8}
s2_for_wx_plot <- s2_interp %>% 
  filter(date>="2020-07-05", date<="2020-09-01") %>% 
  mutate(name=as.factor("Sentinel-2 RGND")) %>% 
  rename(value=rgnd) %>% 
  select(site, date, name, value)

my_wx <- ts_custom_wide %>% 
  filter(platform %in% c("terraSnow")) %>% 
  select(site, date, rgnd, albedo) %>% 
  pivot_longer(-site:-date, values_drop_na =TRUE) %>% 
  pivot_wider() %>% 
  full_join(wx_clean) %>% 
  # # mean of both glaciers
  # group_by(date) %>%
  # select(-site) %>%
  # summarise(across(everything(), list(mean = mean), .names = "{.col}")) %>% 
  pivot_longer(-date:-site, values_drop_na=TRUE) %>%
  bind_rows(s2_for_wx_plot) %>% 
  mutate(name = name %>% fct_relevel("Sentinel-2 RGND", "albedo", "tmax", "tmin", "prcp", "srad") %>% 
           fct_recode("MOD10A1 Albedo"="albedo", "Max. Temp C"="tmax", "Min. Temp C"="tmin", "Precip. (mm)" = "prcp", 
                      "Solar radiation (w/m2)"="srad"),
         site = site %>% fct_recode("Vowell"="vowell", "Catamount"="cat") %>% fct_relevel("Vowell")) %>% 
  drop_na() %>% 
  filter(name!="dayl")

wxplt <- my_wx %>% 
  ggplot(aes(x = date, y = value)) +
  facet_grid(rows = vars(name), cols = vars(site), scales = "free") +
  geom_line(alpha = 1, size=1, color = "#f8766d", data = my_wx %>% filter(name %in% c("Sentinel-2 RGND"))) +
  geom_point(alpha = 0.7, shape = 0, color = "grey30", data = my_wx %>% filter(name %in% c("MOD10A1 Albedo"))) +
  theme_minimal() +
  geom_line(stat="smooth", size=1, data = my_wx %>% filter(name =="MOD10A1 Albedo"), color = "grey50") +
  geom_line(alpha = 0.5, data = my_wx %>% filter(!(name %in% c("Sentinel-2 RGND", "MOD10A1 Albedo")))) +
  # add horiz line to some facets
  geom_hline(aes(yintercept = 0), data = my_wx %>% filter(!(name %in% c("MOD10A1 Albedo", "Solar radiation (w/m2)"))) ) +
  labs(y="",x="Date")
wxplt
```


```{r}
ggsave(here("figs/wx_plot.pdf"),height =9, width = 8 )
```

# Applying G-D melt model, algae melt = 0.13 sqrt(area in cm2/l)

```{r}
melt <- function(rgnd){
  cell_dens <- rgnd * 1727 # my model, in cm2/L
  0.13*sqrt(cell_dens) # G-D melt model, cm water eq./day
}

s2_interp %>% 
  mutate(melt = melt(rgnd)) %>% 
  group_by(site) %>% 
  summarise(melt = sum(melt))

```
46 cm

```{r}
melt(0.05)# 10.2 mm SWE/day
melt(0.039) #9 mm SWE/day
```
1 cm water over the entire bloom polygon area, is a volume of 
```{r}
# in units of m2 and m
1.6e6*0.012 # 19,000 m3/day
2.1e6*0.009 # 19,000 m3 per day
```

